---
title: "Example code for ggplot2"
author: "Kevin Rue-Albrecht"
date: "03/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(patchwork)
library(gridExtra)
```

# Exercise

Use `library(ggplot2)` to load the package.

```{r}
library(ggplot2)
```

## Specify the dataset.

What does `ggplot(diamonds)` do?

```{r}
ggplot(diamonds)
```

> Answer:

## Add the aesthetics.

What does `ggplot(diamonds, aes(x = carat, y = price))` do?

```{r}
ggplot(diamonds, aes(x = carat, y = price)) 
```

> Answer:

## Add geometric objects

- Add data points showing `carat` on the x-axis and `price` on the y-axis.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()
  
```

- Color data points by 'cut'.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point()
  
  
```

- Add a smoothed mean trend line.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point() + geom_smooth()
  
  


```

- Assign that last plot to an object called `obds_diamonds`.

```{r}

obds_diamonds <- ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point() + geom_smooth()
  
obds_diamonds
```


# Exercise

Predict the difference between these two plots

### Plot 1

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
  geom_point() +
  geom_smooth()
```

### Plot 2

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(colour = cut)) +
  geom_smooth()
```

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes( color=color)) +
  facet_grid(cut ~ color)
```



> Answer:

# Exercise

## Trend lines

Using the `ChickWeight` data set:

- Create a scatter plot of weight (y-axis) over time (x-axis).

```{r}
ggplot(ChickWeight, aes(x=Time , y=weight))+
  geom_point()

```

- Color by diet.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ChickWeight, aes(x=Time , y=weight, color = Diet))+
  geom_point()

```

- Add a linear mean trend line for each diet.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ChickWeight, aes(x=Time , y=weight, color = Diet))+
  geom_point() +
  geom_smooth(method="lm")


```

At this point you should be able to visually identify
the diet that leads to the highest mean increase in weight.

> Answer:

- Facet a ribbon of sub-plots, one per diet.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ChickWeight, aes(x=Time , y=weight, color = Diet))+
  geom_point() +
  geom_smooth(method="lm") +
  facet_wrap(~Diet)

```

- Assign that last plot to an object called `obds_chickweight`.

```{r}
obds_chickweight <- ggplot(ChickWeight, aes(x=Time , y=weight, color = Diet))+
  geom_point() +
  geom_smooth(method="lm") +
  facet_wrap(~Diet)
obds_chickweight

```

# Exercise

## Bar plot

- Load the `ggplot2::msleep` data set.

```{r}
ggplot2::msleep

```

- Draw a bar plot of number of observations (i.e., rows) for each taxonomic order
  (i.e, one plot and one bar per taxonomic order).

```{r}
ggplot(ggplot2::msleep, aes(x = order))+
  geom_bar(aes(color=order))

```

- Change the angle and font size of the text for the x-axis ticks (not the axis titles).
  Justify the text of those x-axis ticks as right-aligned.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ggplot2::msleep, aes(x = order))+
  geom_bar(aes(color=order))+
  theme(
    axis.text.x = element_text(angle=90)
  )
theme()                        




```

- Change the value and font size of the title for both x and y axes.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ggplot2::msleep, aes(x = order))+
  geom_bar(aes( fill=genus),colour = "black", linewidth=0.5)+
  theme(
      axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5,size=rel(1.5)))+
  labs(title = "mSleep", x="order", y="observation") +
         theme(axis.title =element_text(size = 16)) +
  theme(plot.title =element_text(hjust = 0.5)) +
```


```{r}
theme(legend.key.size = unit(2, "mm"))
```


```{r}
+
  guides(fill = guide_legend(ncol = 5))

```

- Fill each bar with colors, proportionally to the count of each genus.
  (Copy-paste and extend the code chunk above.)

```{r}
obds_msleep <- ggplot(ggplot2::msleep, aes(x = order))+
  geom_bar(aes( fill=genus),colour = "black", linewidth=0.5)+
  theme(
      axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5,size=rel(1.5)))+
  labs(title = "mSleep", x="order", y="observation") +
         theme(axis.title =element_text(size = 16)) +
  theme(plot.title =element_text(hjust = 0.5)) + 
  guides(fill = guide_legend(ncol = 5))
  obds_msleep






```

From this point onwards, you may need to iteratively resize the text of the ticks and axes for readability.

- Reduce the legend key size.
  (Recommendation: use `unit(2, "mm")`).
  (Copy-paste and extend the code chunk above.)

```{r}








```

- Force the legend to be display in 3 columns.
  (Recommendation: use `guide_legend(...)`).
  (Copy-paste and extend the code chunk above.)

```{r}










```

- Add a contour of thin black lines to the bars.

```{r}










```

- Assign that last plot to an object called `obds_msleep`.

```{r}










```

# Exercise

## Plotting grid

- Collate the plots that we assigend to objects through the day, as a single plot.

  + Plots: `obds_diamonds`, `obds_chickweight`, `obds_msleep`.
  
  + Methods: `cowplot::plot_grid()`, `patchwork`, `gridExtra::grid.arrange()`.

Using `cowplot`.

```{r, fig.height=8, fig.width=8}
library(cowplot)
first_row <- cowplot::plot_grid(obds_chickweight, obds_diamonds, nrows=1
  
  
)

second_row <- cowplot::plot_grid(obds_msleep, nrows=1
  
)
super_plot <- cowplot::plot_grid( first_row, second_row, nrow = 2
  
  
)
super_plot
```

Using `patchwork`.

```{r, fig.height=8, fig.width=8}
library(patchwork)
obds_chickweight + obds_diamonds / obds_msleep

```

Using `gridExtra`.

```{r, fig.height=12, fig.width=12}
library(gridExtra)





```

- Export the new plot in a PDF file, and open it in a PDF viewer (e.g. Adobe Acrobat Reader DC).

You will likely need a few attempts to fine-tune the width and height of the output file.

```{r}
ggsave( 
  filename = "ggplot_graphs.pdf",
  plot=super_plot
  )
```

# Exercise

## Pair programming

- Explore the data set `ggplot2::mpg` and generate the most informative plot that you can!

```{r}
library(patchwork)
obds_chickweight + obds_diamonds
```


```{r}
library(gridExtra)
```


```{r}
gridExtra::grid.arrange(obds_diamonds, obds_chickweight, ncol = 2, nrow = 1) 

```


```{r}
cowplot::plot_grid(obds_chickweight, obds_diamondsI, labels = c(1, 2), ncol = 1, nrow = 2)
3


```
